Re-run, Repeat, Reproduce, Reuse, Replicate:\\Transforming Code into Scientific Contributions

### 2017-10-17 draft.tex
second part of changes. add links in listings

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 1:}} Random walk (R$^0$)}
import random

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 2:}} Re-runnable random walk (R$^1$)}
# Tested with Python 3
import random

@@

The code is running and producing the expected results. The next step is to make sure that you can produce the same output over successive runs of your program. In other words, the next step is to make your program deterministic, producing {\em repeatable} output.  Repeatability is valuable. If a run of the program produces a particularly puzzling result, repeatability allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger. Repeatability is also useful to prove that the program did indeed produce the published results. Repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}. But for sequential programs not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 3:}} Re-runnable, repeatable random walk (R$^2$)}
# Tested with Python 3
import random

@@

% reproducibility implies availability
It is obvious that \emph{reproducibility implies availability}. As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request. While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable. Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare} or \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment. Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}. The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
import sys, subprocess, datetime, random

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
import sys, subprocess, datetime, random

@@

% reproducible, a definition
\emph{Replicability} is the implicit
assumption that any article that does not provide the source code makes: that
the description it provides of the algorithms is sufficiently precise and
complete to re-obtain the results it presents. Therefore, replicating a result means coding a new implementation from scratch that match the conceptual description of the article. This affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way replicating an laboratory experiment can ferret out hidden biases. While every published article
should strive for replicability, it is seldom obtained. In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\

% conceptual description != replicable description
This is because most papers strive to communicate the main ideas behind their contribution is terms as simple and as clear as possible, so that the reader may be able to easily understand them and the results that are presented to him. Trying to ensure replicability in the main text adds a myriad of esoteric details that are not conceptually significant and clutter the explanations. Therefore, unless the writer dedicates an addendum or a section of the supplementary information for technical details specifically aimed at replicability, the information will not be there because there are incentives not to do so.\\

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 6:}} Replicated random walk (R$^5$)}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.2 / NumPy 1.12.0 / macOS 10.12.6 / 64 bits architecture
import random
import numpy as np

>>>>

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 1:}} Random walk (R$^0$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R0.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
import random

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 2:}} Re-runnable random walk (R$^1$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R1.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
# Tested with Python 3
import random

@@

The code is running and producing the expected results. The next step is to make sure that you can produce the same output over successive runs of your program. In other words, the next step is to make your program deterministic, producing {\em repeatable} output.  Repeatability is valuable. If a run of the program produces a particularly puzzling result, repeatability allows you to scrutinize any step of the execution of the program by re-running it again with extraneous prints, or inside a debugger. Repeatability is also useful to prove that the program did indeed produce the published results. Repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}. But for sequential and deterministically parallels programs \citep{Hines:2008, Collange:2015} not depending on analog inputs, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 3:}} Re-runnable, repeatable random walk (R$^2$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R2.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
# Tested with Python 3
import random

@@

% reproducibility implies availability
It is obvious that \emph{reproducibility implies availability}. As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request. While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable. Code \emph{and} input data \emph{and} result data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as \href{https://figshare.com}{Figshare}, \href{https://zenodo.org}{Zenodo}\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment. Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.} or a domain specific database, such as \href{https://senselab.med.yale.edu/modeldb/}{ModelDB} for computational neuroscience. The codes presented in this article are available in the GitHub repository \href{https://github.com/rougier/random-walk}{github.com/rougier/random-walk} and at \href{https://doi.org/10.5281/zenodo.848217}{doi.org/10.5281/zenodo.848217}.\\

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R3.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
import sys, subprocess, datetime, random

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R4.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.2 / macOS 10.12.6 / 64 bits architecture
import sys, subprocess, datetime, random

@@

% reproducible, a definition
\emph{Replicability} is the implicit
assumption that an article that does not provide the source code makes: that
the description it provides of the algorithms is sufficiently precise and
complete to re-obtain the results it presents. Here, replicating implies writing a new code matching the conceptual description of the article, in order to reobtain the same results. Replication affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way that replicating an laboratory experiment in a different laboratory can ferret out subtle biases. While every published article should strive for replicability, it is seldom obtained. In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\

% conceptual description != replicable description
This is because most papers strive to communicate the main ideas behind their contribution in terms as simple and as clear as possible, so that the reader may be able to easily understand them and the results that are presented. Trying to ensure replicability in the main text adds a myriad of esoteric details that are not conceptually significant and clutter the explanations. Therefore, unless the writer dedicates an addendum or a section of the supplementary information for technical details specifically aimed at replicability, the information will not be there because there are incentives not to do so.\\

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\parbox{.8\textwidth}{\textbf{\textsc{Listing 6:}} Replicated random walk (R$^5$)}\parbox{.161\textwidth}{\hfill \href{https://raw.githubusercontent.com/rougier/random-walk/frontiers/random-walk-R5.py}{raw code}, \href{https://doi.org/10.5281/zenodo.848217}{archive}}}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.2 / NumPy 1.12.0 / macOS 10.12.6 / 64 bits architecture
import random
import numpy as np
